\subsection{Motivation}
The kinematics and abundance of satellite galaxies around field dwarfs can be used to constrain
the galaxy-halo connection in low mass regime, as well as the the radial profile of dark matter
halos around dwarf galaxies. Each dwarf system will have very few satellites, if any, above the spectroscopic
selection threshold. Therefore, such constraint can only be done in a purely statistical sense
using a very large sample of dwarfs. It is similar to the idea of using weak gravitational lensing
around dwarf galaxies to constrain the galaxy-halo connection and the dark matter radial profile.
The choice of a dark matter model impacts the dark matter radial profile, and the differences are
expected to be especially pronounced in case of low mass halos.

Mitra et al. (2024) has demonstrated a highly robust Bayesian hierarchical framework to constrain an elaborate
galaxy-halo connection model (the full conditional luminosity function of central and satellite galaxies)
using the projected phase-space distribution of satellite galaxies in the SDSS data, down to Milky Way mass halo
regime. The same principle can be extended to dwarf scales, and the lack of such extension is limited only
by the lack of spectroscopic data for low mass galaxies. With the field gradually moving towards weak lensing
measurements and modelling for low mass halos (Leauthaud et al. 2019), the time is ripe to build complementary
probes of halo mass in the dwarf regime using upcoming and future spectroscopic surveys.

  
\subsection{Target data: high completeness fainter galaxies around field dwarfs}
The requirement for satellite kinematics is a fairly unbiased high-completeness sample of satellite galaxies around
the systems of interest. This would ideally mean a much fainter BGS-like sample, getting spectra of ``all'' galaxies down to
some r-band apparent magnitude threshold (let's assume r=24). To probe dark matter radial profile it is essential to have
small-scale data, and hence mitigating fiber collision using by having multiple passes for some sub-patch of the footprint
will be crucial.


\subsection{Phase-space distribution of satellite galaxies}
  \subsubsection{Projected velocity dispersion}
  
  \subsubsection{Projected velocity kurtosis}

\subsection{Impact of different SIDM models}

\subsection{Halo mass mixing in the data and the model}

\subsection{Projections}
  \subsubsection{Galaxy-halo connection}
  \subsubsection{SIDM parameter inference from DM radial profile constraints}

\subsection{Synergies with dwarf weak lensing}
