The Spec-S5 instrument concept consists of installing new 6-m mirrors on two existing telescopes:  the Mayall Telescope at Kitt Peak National Observatory in Arizona and the Blanco Telescope at Cerro Tololo Inter-American Observatory in Chile.
Each Spec-S5 telescope will consist of a $\sim$13,000 fiber focal plane with a field of view of 2.2 deg$^2$ and a fiber density of $\sim$XXX fibers per arcmin$^2$.
The baseline design deploys 23 spectrographs based on the DESI design, covering a wavelength range from 360 to 980\,nm at resolutions ranging from $R \sim 2000$ at the blue end to $R \sim 5500$ at the red end. 
Each spectrograph will be by $\sim$567 fibers, each with a diameter of 107\,$\mu$m corresponding to an angular size of $\sim$1 arcsec.
A shorter fiber route, improved fiber efficiency, and improved low-noise detectors project a factor of $\sim$30\% improvement in the blue throughput.
Spec-S5 expects to deliver spectra of $\sim$150 million stars over $\sim$20,000\,deg^2 of the high-Galactic-latitude sky.


The Spec-S5 sensitivity is modeled based on the performance of the Dark Energy Spectroscopic Instrument (DESI). The model includes the transmission of the atmosphere (at zenith), the throughput of the telescope (primary and secondary reflective optics), the transmission of the corrector optics, the fiber losses dues to the fiber aperture and throughput (accounting for the shorter fiber length), the spectrograph throughput (assumed to be identical to the DESI 3-arm spectrographs), and the detector quantum efficiency (assuming MAS CCDs for the blue arm). The signal-to-noise ratio computations assume the DESI-measured sky brightness. DESI observations of faint stars show that the radial velocity measurement uncertainty is proportional to the signal-to-noise ratio measured in the z-arm of the spectrograph (i.e., SN\_Z)\footnote{This measurement error will likely be smaller if Spec-S5 incorporates a near-infrared arm in the existing DESI spectrographs, since this will have a higher resolution of $R\approx6500$ (compared with the current Z-arm resolution of $R\approx 5000$).}:

% VRAD_ERR = 10.0**(1.389 - 0.975*np.log10(SN_Z)) - 0.975*np.log10(0.8))
\begin{equation}
\sigma(V_{\rm rad}) = 10^{\rm 1.389 - 0.975log(SN\_Z) - 0.975log(0.8)}
\end{equation}

The model predicts that a $3 \times 1200$~sec observation of a z=22~AB mag star with Spec-S5 will result in a $\sigma(V_{\rm rad})$ of $\approx 6.5~{\rm km,s^{-1}}$.
